### 2013-05-22 inst/examples/BUGS/myers.Rmd
still need to fix up Allen MCMC convergence when underlying model is Myers

```{r allen_pars}
# a bit unfair to start with the correct values, but anyhow...
init_p = p
names(init_p) = c("r0", "K", "theta")
```

@@

```{r allen-jags}
K_prior_p <- c(0.01, 40.0)
logr0_prior_p <- c(-6.0, 6.0)
logtheta_prior_p <- c(-6.0, 6.0)

@@

```{r allen-priors}
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
logtheta_prior <- function(x) dunif(x, -6, 6)
par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
                    logr0 = logr0_prior, logtheta = logtheta_prior,
                    stdQ = stdQ_prior, stdR = stdR_prior)
```

@@

```{r allen-mcmc}
jags.params=c("K","logr0","logtheta","stdQ", "stdR")
jags.inits <- function(){
  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), 
       "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

@@

```{r allen-traces}
jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
allen_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
ggplot(allen_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
```

```{r allen-posteriors}
allen_priors <- ddply(allen_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})

@@

```{r allen-output}
pardist <- acast(allen_posteriors, index ~ variable, subset=.(variable != "deviance") )
pardist[,2] = exp(pardist[,2]) # transform model parameters back first
pardist[,3] = exp(pardist[,3])
bayes_coef <- apply(pardist,2, posterior.mode) 
bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3])) # parameters formatted for f
allen_means <- sapply(x_grid, f, 0, bayes_pars)
```

@@

```{r ricker-jags}
K_prior_p <- c(0.01, 40.0)
logr0_prior_p <- c(-6.0, 6.0)

@@

```{r ricker-priors}
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
                   logr0 = logr0_prior, stdQ = stdQ_prior, stdR = stdR_prior)
```

@@

```{r ricker-mcmc}
# Uniform priors on standard deviation terms
jags.params=c("K","logr0", "stdQ", "stdR")
jags.inits <- function(){
  list("K"=init_p["K"], "logr0"=log(init_p["r0"]), 
       "stdQ"=sqrt(0.05), "stdR"=sqrt(0.1), "x"=y, 
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345) 
jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
                  model.file="ricker.bugs", progress.bar="none")
```

@@

```{r ricker_traces}
jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
ricker_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
  facet_wrap(~ variable, scale="free", ncol=1)
```

```{r ricker_posteriors}
ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})
# plot posterior distributions
ggplot(ricker_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

@@

```{r ricker-output}
ricker_pardist <- acast(ricker_posteriors, index ~ variable, subset=.(variable != "deviance") )
ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates from mode then mean
ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
```

@@

```{r myers-jags}
logr0_prior_p <- c(-6.0, 6.0)
logtheta_prior_p <- c(-6.0, 6.0)
logK_prior_p <- c(-6.0, 6.0)

@@

```{r myers-mcmc}
init_p = c(1, 1, 1)
names(init_p) = c("r0", "theta", "K")
jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
jags.inits <- function(){
  list("logr0"=log(init_p["r0"]), "logtheta" = log(init_p["theta"]),  
       "logK"=log(init_p["K"]), "stdQ"=sqrt(0.05), 
       "stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345)
jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, 
                  n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
                  n.burnin=n.burnin, model.file="myers.bugs", progress.bar="none")

@@

```{r myers-traces}
jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
myers_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
ggplot(myers_posteriors) + geom_line(aes(index, value)) +
  facet_wrap(~ variable, scale="free", ncol=1)

@@

```{r myers-posteriors}
par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})

@@

```{r myers-output}
myers_pardist <- as.matrix(jags_matrix[2:6])
myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
```

@@

models <- melt(models, id="x")
names(models) <- c("x", "method", "value")
```

```{r Figure1}
plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
    geom_line(data=models, aes(x, value, col=method), lwd=1, alpha=0.8) + 
    geom_point(data=obs, aes(x,y), alpha=0.8) + 
    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
    scale_colour_manual(values=cbPalette) 
print(plot_gp)
```

@@

```{r gp-opt, dependson="assemble-models"}
MaxT = 1000
# uses expected values from GP, instead of integrating over posterior
#matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)

@@

```{r mle-opt, dependson="assemble-models"}
matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

@@

```{r allen-opt, dependson="assemble-models"}
allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
opt_allen <- value_iteration(matrices_allen, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r ricker-opt, dependson="assemble-models"}
ricker_f <- function(x, h, p) Ricker(x, h, unname(p[c(2, 1)])) # defined by pdgControl 
matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
opt_ricker <- value_iteration(matrices_ricker, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r myers-opt, dependson="assemble-models"}
myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r assemble-opt, , dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_ricker$D, Allen = opt_allen$D, Myers = myers_alt$D)
colorkey=cbPalette
names(colorkey) = names(OPT) 
```

>>>>

```{r allen_pars}
init_p = c(1, 10, 2)
names(init_p) = c("r0", "K", "theta")
```

@@

```{r allen-model}
K_prior_p <- c(0.01, 40.0)
logr0_prior_p <- c(-6.0, 6.0)
logtheta_prior_p <- c(-6.0, 6.0)

@@

```{r allen-priors, dependson="common-priors"}
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
logtheta_prior <- function(x) dunif(x, -6, 6)
par_priors  <- list(K = K_prior, deviance = function(x) 0 * x, 
                    logr0 = logr0_prior, logtheta = logtheta_prior,
                    stdQ = stdQ_prior, stdR = stdR_prior)
```

@@

```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup")}
jags.params=c("K","logr0","logtheta","stdQ", "stdR")
jags.inits <- function(){
  list("K"=init_p["K"],"logr0"=log(init_p["r0"]),"logtheta"=log(init_p["theta"]), 
       "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

@@

```{r allen-traces, dependson="allen-mcmc"}
jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
allen_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
ggplot(allen_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)
```

```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
allen_priors <- ddply(allen_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})

@@

```{r allen-output, dependson="allen-traces"}
pardist <- acast(allen_posteriors, index ~ variable, subset=.(variable != "deviance") )
pardist[,2] = exp(pardist[,2]) # transform model parameters back first
pardist[,3] = exp(pardist[,3])
bayes_coef <- apply(pardist,2, posterior.mode) 
bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1], bayes_coef[3])) # parameters formatted for f
allen_means <- sapply(x_grid, f, 0, bayes_pars)
```

@@

```{r ricker-model}
K_prior_p <- c(0.01, 40.0)
logr0_prior_p <- c(-6.0, 6.0)

@@

```{r ricker-priors, dependson="common-priors"}
K_prior     <- function(x) dunif(x, K_prior_p[1], K_prior_p[2])
logr0_prior <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
par_priors <- list(K = K_prior, deviance = function(x) 0 * x, 
                   logr0 = logr0_prior, stdQ = stdQ_prior, stdR = stdR_prior)
```

@@

```{r ricker-mcmc, dependson="ricker-model"}
# Uniform priors on standard deviation terms
jags.params=c("K","logr0", "stdQ", "stdR")
jags.inits <- function(){
  list("K"=init_p["K"], "logr0"=log(init_p["r0"]), 
       "stdQ"=sqrt(0.05), "stdR"=sqrt(0.1), "x"=y, 
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345) 
jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
                  model.file="ricker.bugs", progress.bar="none")
```

@@

```{r ricker_traces, dependson="ricker-mcmc"}
jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
ricker_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
ggplot(ricker_posteriors) + geom_line(aes(index, value)) + 
  facet_wrap(~ variable, scale="free", ncol=1)
```

```{r ricker_posteriors, dependson=c("ricker-traces", "ricker-priors")}
ricker_priors <- ddply(ricker_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})
# plot posterior distributions
ggplot(ricker_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=ricker_priors, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

@@

```{r ricker-output, dependson="ricker-traces"}
ricker_pardist <- acast(ricker_posteriors, index ~ variable, subset=.(variable != "deviance") )
ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first
bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates from mode then mean
ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
ricker_means <- sapply(x_grid, Ricker, 0, ricker_bayes_pars[c(1,2)])
```

@@

```{r myers-model}
logr0_prior_p <- c(-6.0, 6.0)
logtheta_prior_p <- c(-6.0, 6.0)
logK_prior_p <- c(-6.0, 6.0)

@@

```{r myers-mcmc, dependson="myers-model"}
init_p = c(p[1], p[3], p[2])
names(init_p) = c("r0", "theta", "K")
jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
jags.inits <- function(){
  list("logr0"=log(init_p["r0"]), "logtheta" = log(init_p["theta"]),  
       "logK"=log(init_p["K"]), "stdQ"=sqrt(0.05), 
       "stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345)
jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, 
                  n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
                  n.burnin=n.burnin, model.file="myers.bugs", progress.bar="none")

@@

```{r myers-traces, dependson="myers-mcmc"}
jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
myers_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")
ggplot(myers_posteriors) + geom_line(aes(index, value)) +
  facet_wrap(~ variable, scale="free", ncol=1)

@@

```{r myers-posteriors, dependson="myers-traces"}
par_prior_curves <- ddply(myers_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})

@@

```{r myers-output, dependson="myers-traces"}
myers_pardist <- as.matrix(jags_matrix[2:6])
myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
```

@@

models <- melt(models, id="x")
names(models) <- c("x", "method", "value")

model_names = c("GP", "True", "MLE", "Ricker", "Allen", "Myers")
colorkey=cbPalette
names(colorkey) = model_names 
```

```{r Figure1, dependson="assemble-models"}
plot_gp <- ggplot(tgp_dat) + geom_ribbon(aes(x,y,ymin=ymin,ymax=ymax), fill="gray80") +
    geom_line(data=models, aes(x, value, col=method), lwd=1, alpha=0.8) + 
    geom_point(data=obs, aes(x,y), alpha=0.8) + 
    xlab(expression(X[t])) + ylab(expression(X[t+1])) +
    scale_colour_manual(values=cbPalette) 
print(plot_gp)
```

## Step-ahead predictors

## Posterior predictive curves

This shows only the mean predictions.  For the Bayesian cases, we can instead loop over the posteriors of the parameters (or samples from the GP posterior) to get the distribution of such curves in each case.

We will need a vector version (`pmin` in place of `min`) of the parametric growth functions that can operate on the posteriors, (with appropriate ordering of parameters as they are in the posterior):

```{r par-fns}
ricker_f <- function(x,h,p){
  sapply(x, function(x){ 
    x <- pmax(0, x-h) 
    pmax(0, x * exp(p[2] * (1 - x / p[1] )) )
  })
}
allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])

```{r Figureb, dependson=c("assemble-models", "par-fns")}
require(MASS)
step_ahead <- function(x, f, p){
  h = 0
  x_predict <- sapply(x, f, h, p)
  n <- length(x_predict) - 1
  y <- c(x[1], x_predict[1:n])
  y
}
step_ahead_posteriors <- function(x){
gp_f_at_obs <- gp_predict(gp, x, burnin=1e4, thin=300)
df_post <- melt(lapply(sample(100), 
  function(i){
    data.frame(time = 1:length(x), stock = x, 
                GP = mvrnorm(1, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),
                True = step_ahead(x,f,p),  
                MLE = step_ahead(x,f,est$p), 
                Allen = step_ahead(x, allen_f, pardist[i,]), 
                Ricker = step_ahead(x, ricker_f, ricker_pardist[i,]), 
                Myers = step_ahead(x, myers_f, myers_pardist[i,]))
  }), id=c("time", "stock"))
}

df_post <- step_ahead_posteriors(x)

ggplot(df_post) + geom_point(aes(time, stock)) + 
  geom_line(aes(time, value, col=variable, group=interaction(L1,variable)), alpha=.1) + 
  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1))) 
```

@@

```{r gp-opt, dependson="gp-output"}
MaxT = 1000
# uses expected values from GP, instead of integrating over posterior
#matrices_gp <- gp_transition_matrix(gp_dat$E_Ef, gp_dat$E_Vf, x_grid, h_grid)

@@

```{r mle-opt, dependson="mle-output"}
matrices_true <- f_transition_matrix(f, p, x_grid, h_grid, sigma_g)
opt_true <- value_iteration(matrices_true, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)

@@

```{r allen-opt, dependson="allen-output"}
allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
matrices_allen <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
opt_allen <- value_iteration(matrices_allen, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r ricker-opt, dependson="ricker-output"}
ricker_f <- function(x, h, p) Ricker(x, h, unname(p[c(2, 1)])) # defined by pdgControl 
matrices_ricker <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
opt_ricker <- value_iteration(matrices_ricker, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r myers-opt, dependson="myers-output"}
myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_ricker$D, Allen = opt_allen$D, Myers = myers_alt$D)
colorkey=cbPalette
names(colorkey) = names(OPT) 
```
